---
title: "Crossing The Sea"
output:
  pdf_document: default
  html_document: default
---
---
title: "CDS help"
output: html_document
---

First we are going to install the needed packages: 

```{r}
library(tidyverse)
library(gganimate)
library(ggthemes)
```

Now we are loading the first dataset that we have tidied in Excel: 

```{r}
library(readr) 
new <- read_csv2("~/Documents/RStudioCultural DataScience /data/Deaths.csv") #loading the dataset
new<-new[,1:5] #grapping the five fist columns
new<-na.omit(new) #removing NA's


```

In the following four graphs we are looking at how many people attempts to cross the eastern, western and central route each month. We have made a graph for each year. 

The graph for 2016:

```{r}
new$month <- factor(new$month, levels = month.name) #arranging months chronologically

new %>% 
  filter(year == "2016") %>% #using filter to select the year
  ggplot(aes(x=month,y=attempted,fill=Route))+ #deciding what should be on the x and y axes and what should be the fill
  geom_bar(stat="identity",position=position_dodge())+ #making it bars
  scale_fill_economist(labels = c("Central", "Eastern", "Western"))+ #choosing design and names for the filling
  theme_economist() + 
  xlab("Months") + #choosing name for x axes
  ylab("Attempted Crossings")+ #choosing name for y axes
  ggtitle("Crossings 2016") #choosing tittle
```

The graph for 2017:

```{r}
new$month <- factor(new$month, levels = month.name) #arranging months chronologically

new %>% 
  filter(year == "2017") %>% #using filter to select the year
  ggplot(aes(x=month,y=attempted,fill=Route))+ #deciding what should be on the x and y axes and what should be the fill
  geom_bar(stat="identity",position=position_dodge())+ #making it bars
  scale_fill_economist(labels = c("Central", "Eastern", "Western"))+ #choosing design and names for the filling
  theme_economist() + 
  xlab("Months") + #choosing name for x axes
  ylab("Attempted Crossings")+ #choosing name for y axes
  ggtitle("Crossings 2017") #choosing tittle
```

The graph for 2018:

```{r}
new$month <- factor(new$month, levels = month.name) #arranging months chronologically

new %>% 
  filter(year == "2018") %>% #using filter to select the year
  ggplot(aes(x=month,y=attempted,fill=Route))+ #deciding what should be on the x and y axes and what should be the fill
  geom_bar(stat="identity",position=position_dodge())+ #making it bars
  scale_fill_economist(labels = c("Central", "Eastern", "Western"))+ #choosing design and names for the filling
  theme_economist() + 
  xlab("Months") + #choosing name for x axes
  ylab("Attempted Crossings")+ #choosing name for y axes
  ggtitle("Crossings 2018") #choosing tittle
```

The graph for 2019:

```{r}
new$month <- factor(new$month, levels = month.name) #arranging months chronologically

new %>% 
  filter(year == "2019") %>% #using filter to select the year
  ggplot(aes(x=month,y=attempted,fill=Route))+ #deciding what should be on the x and y axes and what should be the fill
  geom_bar(stat="identity",position=position_dodge())+ #making it bars
  scale_fill_economist(labels = c("Central", "Eastern", "Western"))+ #choosing design and names for the filling
  theme_economist() + 
  xlab("Months") + #choosing name for x axes
  ylab("Attempted Crossings")+ #choosing name for y axes
  ggtitle("Crossings 2019") #choosing tittle
```

Now we are going to read a new tidied version of the sama dataset. We have made a new dataset so that it was easier for us to look at all of the years in two different graphs. With these we can see the attempted crossings per route in one and how many died on another graph. 

```{r}
library(readr)
data <- read_csv2("~/Documents/RStudioCultural DataScience /data/johan_arrivals.csv")


data<-data[,1:4] #grapping the four fist columns

data<-na.omit(data) #removing NA's

data$year<-as.factor(data$year) #making year a factor instead of numeric.
```


```{r}
ggplot(data=data, aes(x=year, y=deaths,fill = route)) + #choosing what should be shown in the graph
  ggtitle("Deaths in the Mediterrenean") + #choosing title
  geom_bar(stat="identity", position=position_dodge()) + 
  theme_economist()+scale_fill_economist()+ #choosing design
  ylab("Deaths") #name for y axis

ggplot(data=data, aes(x=year, y=total_attempts,fill = route)) + #choosing what should be shown in the graph
  ggtitle("Attempted Crossings") + #choosing title
  geom_bar(stat="identity", position=position_dodge()) + 
  theme_economist()+scale_fill_economist() + #choosing design
  ylab("Total Attempts") #choosing name for y axis
```


